Data Analysis Methods for the Detection of the Epoch of Reionization

### 2017-11-22 PAPER.tex
made jonnie edits until section 3.2

To capture the full statistical likelihood of signal loss, one requires a quick way to generate many realizations of simulated $21$ cm signal visibilities. Here we use the same method used in P64, where mock gaussian noise visibilities (mock EoR signals) are filtered in fringe-rate space to retain only ``sky-like" time-modes. This signal is then added to the data.  \footnote{One specific change from P64 is that we add this simulated signal into the chain before the final fringe-rate filter is applied to the data. Previously, the addition was done after that final fringe-rate filter step.  This change results in an increased estimate of signal loss, likely due to the use of the fringe-rate filter as a simulator. Whether the loss would be estimated at the same level if different $21$ cm simulations were used remains an open question.}

@@

\noindent where the data-only term (first term) and EoR-only term (last term) are approximately diagonal matrices. The middle two terms, however, have a larger amount of power in their off-diagonal terms relative to their diagonal elements, since $\textbf{e}$ and $\textbf{x}$ are statistically unrelated. Upon inversion, these off-diagonal terms become negative, driving the cross-terms in Equation \ref{eq:pout_expand} to have negative power. This effect is explained in more depth in \citet{switzer_et_al2015}.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_transfercurve.png}
	\caption{Signal loss transfer functions showing the relationship of $P_{in}$ and $P_{out}$, as defined by Equations \eqref{eq:Pin} and \eqref{eq:sigloss}. Kernel density estimations of the power spectrum transfer functions are shown as colored heat-maps for the cases of inverse covariance weighted PAPER-64 data (left) and unweighted data (right). The solid black diagonal line marks a perfect unity mapping, and the solid grey horizontal line denotes the peak of $\hat{\textbf{p}}$, the data distribution. From these plots, it is clear that inverse covariance weighting results in $\sim3$ orders of magnitude of signal loss for power spectrum values above $\sim$$2 \times 10^{4}$ $mK^{2} \, (h^{-1} \, Mpc)^{3}$, whereas the unweighted case does not exhibit loss.}
	\label{fig:sigloss_transfercurve}
\end{figure*}

@@

\noindent we fit a polynomial to each, representing mean signal loss functions. We then account for the second component of scatter as described previously by de-convolving out $P_{out, unweighted}$ from $P_{out, weighted}$ for each $P_{in}$ value:

@@

We note that for PAPER-64, $\hat{\textbf{p}}$ is sometimes narrowly peaked at positive power spectrum values (signifying detections of power), and sometimes centered around zero, representing measurements that are consistent with zero.

@@

This results in a final $P_{in}$ distribution --- the distribution of our data as seen through the signal loss lens. We compute power spectrum points from the peak of the histograms, and power spectrum errors from $95\%$ confidence intervals.

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps2_data_nosigloss.png}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data.png}
	\caption{Full inverse covariance weighted power spectrum of PAPER-64 data (positive (black) and negative (grey) points, with $2\sigma$ error bars) before signal loss correction (left) and after (right). The dashed blue line is the unweighted power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical noise level prediction based on observational parameters.}
	\label{fig:ps2_data}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_modeloop.png}
	\caption{Power spectra $2\sigma$ upper limits for $k=0.393$ for fringe-rate filtered PAPER-64 data. Values are shown before (grey) and after (black) signal loss correction as a function of number of eigenmodes of $\hat{\textbf{C}}$ that are down-weighted. This regularization knob is tuned from $0$ modes (i.e. unweighted) to $21$ modes (i.e. full inverse covariance weighting). Over $\sim3$ orders of magnitude of signal loss results when using inverse covariance weighting. Also plotted for comparison are $2\sigma$ power spectrum upper limits for the unweighted case (dashed blue), inverse variance weighted case (red), and added-identity case (cyan). All three regularizations shown (black, red, cyan) perform similarly at their best. Finally, a theoretical prediction for noise ($2\sigma$ error) is plotted as solid green.}
	\label{fig:sigloss_modeloop}
\end{figure*}

@@

In this section we have shown three simple ways of regularizing $\hat{\textbf{C}}$ to minimize signal loss using PAPER-64 data. There are many other weighting schemes that we leave for consideration in future work. For example, one could estimate $\hat{\textbf{C}}$ using information from different subsets of baselines. For redundant arrays this could mean calculating $\hat{\textbf{C}}$ from a different but similar baseline type, such as the $\sim30$ m diagonal PAPER baselines (instead of the horizontal E/W ones). Alternately, covariances could be estimated from all other baselines except the two being cross-multiplied when forming a power spectrum estimate. This method was used in \citet{parsons_et_al2014} in order to avoid suppressing the $21$ cm signal, and it's worth noting that the PAPER-32 results are likely safe from the issue of signal loss underestimation because of this very reason.

>>>>

To capture the full statistical likelihood of signal loss, one requires a quick way to generate many realizations of simulated $21$ cm signal visibilities. Here we use the same method used in P64, where mock Gaussian noise visibilities (mock EoR signals) are filtered in fringe-rate space to retain only ``sky-like" time-modes. This signal is then added to the data.  \footnote{One specific change from P64 is that we add this simulated signal into the analysis pipeline before the final fringe-rate filter is applied to the data. Previously, the addition was done after that final fringe-rate filter step.  This change results in an increased estimate of signal loss, likely due to the use of the fringe-rate filter as a simulator. Whether the loss would be estimated at the same level if different $21$ cm simulations were used remains an open question.}

@@

\noindent where the data-only term (first term) and EoR-only term (last term) are approximately diagonal matrices. The middle two terms, however, have a larger amount of power in their off-diagonal terms relative to their diagonal elements, since $\textbf{e}$ and $\textbf{x}$ are statistically unrelated. Upon inversion, these off-diagonal terms become negative, driving the cross-terms in Equation \ref{eq:pout_expand} to have negative power. \cc{someone help me write this more elegantly...}This effect is explained in more depth in \citet{switzer_et_al2015}.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_transfercurve.png}
	\caption{Signal loss transfer functions showing the relationship of $P_{in}$ and $P_{out}$, as defined by Equations \eqref{eq:Pin} and \eqref{eq:sigloss}. Kernel density estimations of the power spectrum transfer functions are shown as colored heat-maps for the cases of inverse covariance weighted PAPER-64 data (left) and unweighted data (right). The solid black diagonal line marks a perfect unity mapping, and the solid grey horizontal line denotes the peak of $\hat{\textbf{p}}$, the data distribution. From these plots, it is clear that inverse covariance weighting results in $\sim3$ orders of magnitude of signal loss for power spectrum values above $\sim$$2 \times 10^{4}$ $mK^{2} \, (h^{-1} \, Mpc)^{3}$, whereas the unweighted case does not exhibit loss. The peculiar feature of `signal gain' at low injection levels is explained in the text.}
	\label{fig:sigloss_transfercurve}
\end{figure*}

@@

\noindent we fit an $8$th order polynomial to each, representing mean signal loss functions. We then account for the second component of scatter as described previously by de-convolving out $P_{out, unweighted}$ from $P_{out, weighted}$ for each $P_{in}$ value:

@@

We note that for PAPER-64, depending on the $k$-value, $\hat{\textbf{p}}$ is sometimes narrowly peaked at positive power spectrum values (signifying detections of power) and sometimes centered around zero, representing measurements that are consistent with zero.

@@

This results in a final $P_{in}$ distribution --- the distribution of our data as seen through the signal loss lens. We compute power spectrum points from the peak of the histograms, and power spectrum errors from $95\%$ confidence intervals (recall that the spread in these probability density functions mostly comes from bootstrapping the data).

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps2_data_nosigloss.png}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data.png}
	\caption{Full inverse covariance weighted power spectrum of PAPER-64 data (positive (black) and negative (grey) points, with $2\sigma$ error bars) before signal loss correction (left) and after (right). The dashed blue line is the unweighted power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical noise level prediction based on observational parameters, whose calculation is detailed in Section \ref{sec:PSSense}.}
	\label{fig:ps2_data}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_modeloop.png}
	\caption{Power spectra $2\sigma$ upper limits for $k=0.393$ h Mpc$^{-1}$ for fringe-rate filtered PAPER-64 data. Values are shown before (grey) and after (black) signal loss correction as a function of number of eigenmodes of $\hat{\textbf{C}}$ that are down-weighted. This regularization knob is tuned from $0$ modes (i.e. unweighted) to $21$ modes (i.e. full inverse covariance weighting). Over $\sim3$ orders of magnitude of signal loss results when using inverse covariance weighting. Also plotted for comparison are $2\sigma$ power spectrum upper limits for the unweighted case (dashed blue), inverse variance weighted case (red), and added-identity case (cyan). All three regularizations shown (black, red, cyan) perform similarly (when $\sim$$3$ modes are down-weighted for the black curve). Finally, a theoretical prediction for noise ($2\sigma$ error) is plotted as solid green.}
	\label{fig:sigloss_modeloop}
\end{figure*}

@@

In this section we have shown three simple ways of regularizing $\hat{\textbf{C}}$ to minimize signal loss using PAPER-64 data. There are many other weighting schemes that we leave for consideration in future work. For example, one could estimate $\hat{\textbf{C}}$ using information from different subsets of baselines. For redundant arrays this could mean calculating $\hat{\textbf{C}}$ from a different but similar baseline type, such as the $\sim30$ m diagonal PAPER baselines (instead of the horizontal E/W ones). Alternately, covariances could be estimated from all other baselines except the two being cross-multiplied when forming a power spectrum estimate. This method was used in \citet{parsons_et_al2014} in order to avoid suppressing the $21$ cm signal, and it's worth noting that the PAPER-32 results are likely safe from the issue of signal loss underestimation because of this very reason (however, they are affected by the error estimation issues described in Section \ref{sec:Error}, so we also regard those results as suspect and superseded by those of \cc{cite kolopanis}).
